---
title: "Predicting Delayed Flights"
output:
  html_document:
    df_print: paged
  word_document: default
---
```{r}
library(readr)
library(caret)
library(ISLR)
#install.packages("e1071") #install first
library(e1071)  
library(cluster)
library(dplyr)
library(pROC)
```

# Read the data and Summary

```{r}
MyData <- read.csv("FlightDelays.csv")
summary(MyData)
```

# Clean the data

```{r}
MyData <- MyData[,c(-3,-5,-6,-7,-9,-11,-12)]
str(MyData)
head(MyData)
set.seed(123)
```

# Week and Time Variables to be recorded as Factors

```{r}
MyData$DAY_WEEK <- as.factor(MyData$DAY_WEEK)
levels(MyData$DAY_WEEK)
MyData$CRS_DEP_TIME <- as.factor(MyData$CRS_DEP_TIME)
levels(MyData$CRS_DEP_TIME)
```

# The outcome variable is whether the flight was delayed, and thus it has two classes (1 = delayed and 0 = on time)

```{r}
MyData$Flight.Status <- factor(MyData$Flight.Status,levels = c("delayed","ontime"),labels = c(0,1))
```

# Divide the data into training and Validation

```{r}
# 60% reserved for Training
Train_Index <- createDataPartition(MyData$Flight.Status, p=0.6, list=FALSE)
Training <- MyData[Train_Index,]
# Validation is the rest 40%
Valid_Data  <- MyData[-Train_Index,]
```

# Run the Naive Bayes model to predict whether the flight is delayed or not. Use only categorical variables for the predictor variables. 

```{r}
nb_model <-naiveBayes(Training$Flight.Status~CARRIER+DEST+ORIGIN+DAY_WEEK+CRS_DEP_TIME, data = Training)
nb_model
```


# Output the confusion matrix and ROC for the validation data

```{r}
# Predicting the delayed status on Validation dataSet
Predicted_Valid_labels <-predict(nb_model,Valid_Data)
library("gmodels")
# Show the confusion matrix of the classifier
CrossTable(x=Valid_Data$Flight.Status,y=Predicted_Valid_labels, prop.chisq = FALSE) 
nb_model <- naiveBayes(Training$Flight.Status~CARRIER+DEST+ORIGIN+DAY_WEEK+CRS_DEP_TIME,data = Training)
#Make predictions and return probability of each class
Predicted_Valid_labels <-predict(nb_model,Valid_Data, type = "raw")
#show the first few values 
head(Predicted_Valid_labels)
```

# ROC Curve for Validation Data Set

```{r}
roc(Valid_Data$Flight.Status, Predicted_Valid_labels[,2])
plot.roc(Valid_Data$Flight.Status,Predicted_Valid_labels[,2])
```

# Output both a counts table and a proportion table outlining how many and what proportion of flights were delayed and on-time at each of the three airports.

```{r}
# Counts Table
table(MyData$Flight.Status, MyData$DEST)

# Proportion Table
prop.table(table(MyData$Flight.Status , MyData$DEST))
```


